---
title: "Homework 2"
author: "Philip Kim"
date: "10/1/2021"
output: 
  github_document:
    toc: true
    toc_depth: 6
---

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

## Read/Clean Mr. Trash Wheel Dataset

```{r}
trashwheel_df = read_excel(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = 1,
  range = "A2:N406") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

## Clean Precipitation Data for 2018 and 2019

### Precipitation Data for 2018

```{r}
precipitation_2018 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = 5,
  skip = 1) %>%
  drop_na(Month) %>% 
  mutate(year = "2018")
```


### Precipitation Data for 2019

```{r}
precipitation_2019 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = 4,
  skip = 1) %>% 
  drop_na() %>% 
  mutate(year = "2019")
```


### Combining Precipitation Data

```{r}
precipitation_combined = 
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  janitor::clean_names() %>% 
  relocate(year) %>% 
  rename(total_in = total) %>% 
  mutate(month = month.name[month])
```


## Data Observations Paragraph



# Problem 2

## Clean pols-month data 

```{r, message = FALSE}
politician_df = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    month = month.name[month],
    president = case_when(prez_gop == "1" ~ "gop", prez_dem == "1" ~ "dem")) %>%  
  select(-prez_gop, -prez_dem, -day)
```


## Clean snp data

```{r message = FALSE}
snp_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, c("month", "day", "year"), convert = TRUE) %>% 
  mutate(
    month = factor(month.name[month], levels = month.name),
    year = if_else(year > 15, year + 1900, year + 2000)) %>% 
  arrange(year, month) %>% 
  select(year, month, close)
```


## Clean unemployment data

```{r message = FALSE}
unemployment_df = read_csv("./data/fivethirtyeight_datasets/unemployment.csv")

unemployment_tidy_df = 
  pivot_longer(
    unemployment_df,
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_perc") %>% 
  janitor::clean_names() %>% 
  mutate(month = match(month, month.abb),
         month = month.name[month])
```


## Join Datasets

```{r}
snp_pols_df = 
  left_join(politician_df, snp_df, by = c("year", "month"))

snp_pols_unemp_df =
  left_join(snp_pols_df, unemployment_tidy_df, by = c("year", "month"))
```

## Dataset Description



# Problem 3

## Tidying popularity of baby names dataset

```{r message = FALSE}
baby_name_df = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_lower(gender),
    childs_first_name = str_to_lower(childs_first_name),
    ethnicity = str_to_lower(ethnicity),
    ethnicity = recode(ethnicity,
                   "asian and paci" = "asian and pacific islander",
                   "black non hisp" = "black non hispanic",
                   "white non hisp" = "white non hispanic")) %>% 
  distinct()
```


## Popularity of name "Olivia"

```{r}
olivia_pop_df = baby_name_df %>% 
  filter(childs_first_name == "olivia") %>% 
  select(-gender, -childs_first_name, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>% 
  relocate(ethnicity, "2011", "2012", "2013", "2014", "2015")
```


## Male name popularity

```{r}
boy_pop_df = baby_name_df %>% 
  filter(
    gender == "male",
    rank == "1") %>% 
  select(-count, -gender, -rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>% 
  relocate(ethnicity, "2011", "2012", "2013", "2014", "2015")
```


## 2016 White non-hispanic male children plot 

```{r}
male_2016_df = baby_name_df %>% 
  filter(gender == "male",
         ethnicity == "white non hispanic",
         year_of_birth == "2016")

ggplot(male_2016_df, aes(x = rank, y = count)) + 
  labs(
    title = "Name Popularity in Male, White \nNon-Hispanic Children Born in 2016",
    x = "Popularity Rank of Name", 
    y = "Number of Children with Respective Name") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_point()
```

