---
title: "Homework 2"
author: "Philip Kim"
date: "10/1/2021"
output: 
  github_document:
    toc: true
    toc_depth: 6
---

```{r, message = FALSE}
library(tidyverse)
library(readxl)
library(knitr)
```

# Problem 1

## Read/Clean Mr. Trash Wheel Dataset

```{r}
trashwheel_df = read_excel(
  "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = 1,
  range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  drop_na(date) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

## Clean Precipitation Data for 2018 and 2019

### Precipitation Data for 2018

```{r}
precipitation_2018 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = 7,
  skip = 1) %>%
  drop_na(Month) %>% 
  mutate(year = "2018")

kable(precipitation_2018, caption = "Precipitation 2018")
```


### Precipitation Data for 2019

```{r}
precipitation_2019 = read_excel(
  "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = 6,
  skip = 1) %>% 
  drop_na() %>% 
  mutate(year = "2019")

kable(precipitation_2019, caption = "Precipitation 2019")
```


### Combining Precipitation Data

```{r}
precipitation_combined = 
  bind_rows(precipitation_2018, precipitation_2019) %>% 
  janitor::clean_names() %>% 
  relocate(year) %>% 
  rename(total_prcp_in = total) %>% 
  mutate(month = month.name[month])

kable(precipitation_combined, caption = "Precipitation (2018 and 2019)")
```


## Data Observations Paragraph

The Mr. Trash Wheel dataset ("trashwheel_df") contains `r nrow(trashwheel_df)` rows and `r ncol(trashwheel_df)` columns. The different `dumpster`s are indicated by numbers and the `date` corresponds to the date of trash collection. Within these, the dataset includes dumpster-specific data such as how much trash it has intercepted which is denoted by the weight of the trash or, `weight_tons`. Accordingly, the types of trash that are intercepted are also specified which includes `plastic_bottles`, `cigarette_butts`, `chip_bags`, and many more. The dumpster that had collected the most amount of trash by weight was dumpster 62 with a collection of   `r max(pull(trashwheel_df, weight_tons))` tons. In contrast, dumpster 407 had the least collection with a total of `r min(pull(trashwheel_df, weight_tons))` tons. Accordingly, the median number of sports balls in a dumpster in 2019 was `r median(pull(filter(trashwheel_df, year == 2019), sports_balls))`.

As for the combined precipitation data ("precipitation_combined"), it looks at the total amount of precipitation per `month` by `year`. This dataset has `r nrow(precipitation_combined)` rows and `r ncol(precipitation_combined)` columns. The total amount of precipitation, in inches, for the year of 2018 was `r sum(pull(precipitation_2018, Total))` inches.


# Problem 2

## Clean pols-month data 

```{r, message = FALSE}
politician_df = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    month = month.name[month],
    president = case_when(prez_gop == "1" ~ "gop", prez_dem == "1" ~ "dem")) %>%  
  select(-prez_gop, -prez_dem, -day)
```


## Clean snp data

```{r message = FALSE}
snp_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, c("month", "day", "year"), convert = TRUE) %>% 
  mutate(
    month = factor(month.name[month], levels = month.name),
    year = if_else(year > 15, year + 1900, year + 2000)) %>% 
  arrange(year, month) %>% 
  select(year, month, close)
```


## Clean unemployment data

```{r message = FALSE}
unemployment_df = read_csv("./data/fivethirtyeight_datasets/unemployment.csv")

unemployment_tidy_df = 
  pivot_longer(
    unemployment_df,
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_perc") %>% 
  janitor::clean_names() %>% 
  mutate(month = match(month, month.abb),
         month = month.name[month])
```


## Join Datasets

```{r}
snp_pols_df = 
  left_join(politician_df, snp_df, by = c("year", "month"))

snp_pols_unemp_df =
  left_join(snp_pols_df, unemployment_tidy_df, by = c("year", "month"))
```

## Dataset Description

The politician data ("politician_df") contains `r nrow(politician_df)` rows and `r ncol(politician_df)` columns. This dataset contains information on how many national politicians were democratic or republican on the given date. For example, it lists how many national governors were republican or democratic on a certain `month` of a particular `year` through `gov_gop` and `gov_dem`, respectively; and namely, the `president` indicates whether the president was republic or democratic via values of `gop` and `dem`, respectively.

The Standard & Poor's stock market index data ("snp_df") contains `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. This dataset contains the dates, in `year` and `month`, and `close` which are the closing values of the Standard & Poor's stock index, whcih is an indicator of the stock market at the given time. 

The unemployment dataset ("unemployment_tidy_df") contains `r nrow(unemployment_tidy_df)` rows and
`r ncol(unemployment_tidy_df)` columns. This dataset contains the unemployment percentage at any given time, which is denoted by `unemployment_perc`. 

The resulting data, which is a combination of all the aforementioned datasets, has `r nrow(snp_pols_unemp_df)` rows and `r ncol(snp_pols_unemp_df)` columns. This dataset gives valuable information of the state of the economy given the political standings of the government. It contains information from January 1947 to June 2015 and between those dates, we see that the highest unemployment rate is `r max(pull(snp_pols_unemp_df, unemployment_perc), na.rm = TRUE)`% on November/December of 1982 under the presidency of a republican president. To contrast, the lowest unemployment rate (`r min(pull(snp_pols_unemp_df, unemployment_perc), na.rm = TRUE)`%) was in May/June of 1953 also under the presidency of a republican president.


# Problem 3

## Tidying popularity of baby names dataset

```{r message = FALSE}
baby_name_df = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_lower(gender),
    childs_first_name = str_to_lower(childs_first_name),
    ethnicity = str_to_lower(ethnicity),
    ethnicity = recode(ethnicity,
                   "asian and paci" = "asian and pacific islander",
                   "black non hisp" = "black non hispanic",
                   "white non hisp" = "white non hispanic")) %>% 
  distinct()
```


## Popularity of name "Olivia"

```{r}
olivia_pop_df = baby_name_df %>% 
  filter(childs_first_name == "olivia") %>% 
  select(-gender, -childs_first_name, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank") %>% 
  relocate(ethnicity, "2011", "2012", "2013", "2014", "2015")

kable(olivia_pop_df, caption = "Popularity rank of name Olivia over time")
```


## Male name popularity

```{r}
boy_pop_df = baby_name_df %>% 
  filter(
    gender == "male",
    rank == "1") %>% 
  select(-count, -gender, -rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name") %>% 
  relocate(ethnicity, "2011", "2012", "2013", "2014", "2015")

kable(boy_pop_df, caption = "Most popular name among male children over time")
```


## 2016 white non-hispanic male children plot 

```{r}
male_2016_df = baby_name_df %>% 
  filter(
    gender == "male",
    ethnicity == "white non hispanic",
    year_of_birth == "2016")

ggplot(male_2016_df, aes(x = rank, y = count)) + 
  labs(
    title = "Name Popularity in Male, White \nNon-Hispanic Children Born in 2016",
    x = "Popularity Rank of Name", 
    y = "Number of Children with Respective Name") + 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_point()
```

